---
title: "course5Project1"
author: "ChanterJohnDavid"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Repo

    Valid GitHub URL
    At least one commit beyond the original fork
    Valid SHA-1
    SHA-1 corresponds to a specific commit

Commit containing full submission

    Code for reading in the dataset and/or processing the data
    Histogram of the total number of steps taken each day
    Mean and median number of steps taken each day
    Time series plot of the average number of steps taken
    The 5-minute interval that, on average, contains the maximum number of steps
    Code to describe and show a strategy for imputing missing data
    Histogram of the total number of steps taken each day after missing values are imputed
    Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
    All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
    
Imputing Data : I used is.na to detect NA values and I set them to 0 since 0 will not affect my calculations.  The plots did not make any sense when I did not set the NA values to 0.  All the plots save one show data that used my imputing strategy.  The plot that did not use the imputing strategy is clearly marked and is the Average Steps Taken per Five Minute interval across each day.

All the code is here and is shown

The Panel Plots are the paneled box plots at the end of the html output/
```{r}
library("dplyr")
library("ggplot2")
library("mlbench")
library("caret")
library("chron")



activityDF <- read.csv(file="/Users/douglasashworth/Documents/Coursera/DataScience/Course5/activity.csv", stringsAsFactors = FALSE)

activityDF$steps[is.na(activityDF$steps)] <- 0

summary(activityDF)
str(activityDF)
names(activityDF)

print(head(activityDF))
print(summary(activityDF))
print(str(activityDF))

hist(activityDF$steps, col="green")
maxCnt <- 2400/5 
totalNumDays <- 0
cntr <- 1

while (cntr <= nrow(activityDF)){
  
  fiveMinCntr <- 0
  while (fiveMinCntr <= 2355){
     
     cntr <- cntr + 1
     fiveMinCntr <- fiveMinCntr + 5
     
  }
  
  totalNumDays <- totalNumDays + 1
  
} 
 
fiveMinMatrix = matrix(nrow=473,ncol=totalNumDays)

print(dim(fiveMinMatrix))
 avStepsPerDay <- c()
 medianStepsPerDay <- c()
 maxStepsIntervalPerDay <- c()
 maxStepPerDay <- c()
 maxInterval <- c()
 dailySteps <- c()
 totalNumStepsPerDay <- c()
 
 cntr <- 1
 
 print(maxCnt)
 fiveMinCntr <- 0
 #print(nrow(activityDF))
 print(activityDF$steps[0])
 dayCntr <- 0
 dCntr <- 1
 weekDayStepsAry <- c()
 weekendStepsAry <- c()
 

while (cntr <= nrow(activityDF)){
  
   fiveMinCntr <- 0
   tmpStepSum <- 0.0
   tmpCntr <- 0
   maxStep <- 0
   tmpSteps <- 0
   rowCntr <- 0
   tmpStepAry <- c()
   currentDate <- activityDF$date[cntr]
   
   if (!is.weekend(currentDate)){
     wkdayFlag <- 1
     wkendFlag <- 0
   }else{
     wkendFlag <- 1
     wkdayFlag <- 0
   }
   
   fCntr <- 1
   while (fiveMinCntr <= 2355){
     
     tmpSteps <-  as.numeric(activityDF$steps[cntr])
     tmpStepAry <- c(tmpStepAry,tmpSteps)

     fiveMinMatrix[fCntr, dCntr] <- tmpSteps
     
     cntr <- cntr + 1
     fiveMinCntr <- fiveMinCntr + 5
     fCntr <- fCntr + 1
     
   }
   
   meanSteps <- mean(tmpStepAry)
   medianSteps <- median(sort(tmpStepAry))
   maxSteps <- max(tmpStepAry)
   maxInterval <- which.max(tmpStepAry)*5
   totalNumSteps <- sum(tmpStepAry)
   
   avStepsPerDay <- c(avStepsPerDay,meanSteps)
   medianStepsPerDay <- c(medianStepsPerDay,medianSteps) 
   maxStepPerDay <- c(maxStepPerDay,maxSteps)
   maxStepsIntervalPerDay <- c(maxStepsIntervalPerDay,maxInterval)
   totalNumStepsPerDay <- c(totalNumStepsPerDay,totalNumSteps)
   
   if (wkdayFlag == 1){
     weekDayStepsAry <- c(weekDayStepsAry,meanSteps)
   }else{
     weekendStepsAry <- c(weekendStepsAry,meanSteps)
   }
   
   dayCntr <- dayCntr + 1
   dCntr <- dCntr + 1
    
}  
print(dim(fiveMinMatrix))

aveFiveMinStepsAryNA <- c()

for (m in 1:nrow(fiveMinMatrix)){
  
  tmpSum <- 0.0
  tmpAve <- 0.0
  
  tmpSum <- sum(fiveMinMatrix[m,])
  tmpAve <- tmpSum/38
  
  aveFiveMinStepsAryNA <- c(aveFiveMinStepsAryNA,tmpAve)

}


aveFiveMinStepsAry <- c()

fiveMinMatrix[is.na(fiveMinMatrix)] <- 0

for (m in 1:nrow(fiveMinMatrix)){
  
  tmpSum <- 0.0
  tmpAve <- 0.0
  
  tmpSum <- sum(fiveMinMatrix[m,])
  tmpAve <- tmpSum/38
  
  aveFiveMinStepsAry <- c(aveFiveMinStepsAry,tmpAve)

}

print(length(aveFiveMinStepsAry))

hist(totalNumStepsPerDay, col="blue",main="Total Number of Steps per Day")

day <- 1:dayCntr
fvNA <- 1:length(aveFiveMinStepsAryNA)
qplot(fvNA,aveFiveMinStepsAryNA,geom = c("point","line"), xlab = "Five Minute Interval", ylab = "Average Number of Steps",  main = "Average Number of Steps per Five Minute Interval Across Each Day No Imputing Strategy")
fv <- 1:length(aveFiveMinStepsAry)
qplot(fv,aveFiveMinStepsAry,geom = c("point","line"), xlab = "Five Minute Interval", ylab = "Average Number of Steps",  main = "Average Number of Steps per Five Minute Interval Across Each Day using Imputing Strategy")
qplot(day,totalNumStepsPerDay,geom = c("point","line"), main = "Total Number of Steps per Day")
qplot(day,avStepsPerDay,geom = c("point", "line"),main="Average Number of Steps per Day")
qplot(day,medianStepsPerDay,geom = c("point","line"),main = "Median Number of Steps per Day")
qplot(day,maxStepPerDay,geom = c("point","line"), xlab = "Day",ylab = "Maximum Number of Steps",main = "Maxiumun Steps Per Day")
qplot(day,maxStepsIntervalPerDay,geom = c("point","line"), xlab = "Day",ylab = "Maximum Five Minute Interval",main = "Maximun Step Interval Per Day")

wkday <- 1:length(weekDayStepsAry)
wkendDay <- 1:length(weekendStepsAry)
par(mfrow = c(1,2))
plot(wkday,weekDayStepsAry)
plot(wkendDay,weekendStepsAry)

par(mfrow = c(1,2))
boxplot(weekDayStepsAry, main="Average Weekday Steps")
boxplot(weekendStepsAry, main="Average Weekend Steps")
#plot(maxStepsIntervalPerDay)

```

## Including Plots

You can also embed plots, for example:

```{}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
